\documentclass[prodmode,acmtecs]{acmsmall}
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}



\begin{document}
\markboth{B D Ball}{Technologies for Real-Time Collaboration in BioInformatics with Synchronized Web Browsers}

\title{Technologies for Real-Time Collaboration using Synchronized Web Browsers in BioInformatics} % title
\author{Brendan D Ball
\affil{University of Cape Town}}
\date{April 2015}

\begin{abstract}
This is an abstract
\end{abstract}

\keywords{WebRTC, WebSocket, Real-time collaboration, Bioinformatics}
\maketitle
\section{Introduction}

Bioinformatics has become a fast growing field both as a result of advancements in technology and advancements in experimental biological studies. This is evident when looking at the Human Genome Project. These advancements have created a need for collaboration amongst scientists, often working together from across the globe and often across disciplines \cite{li2013ngs,lee2007facilitating,tsiliki2014datamining}. One solution to this problem has been the SEQanswers project. This project tries to improve knowledge sharing by means of an open forum where scientists are able to have discussions specific to genomics \cite{li2012seqanswers}. Another successful community-driven project is the NGS WikiBook, which as the name implies, is a wiki containing lots of information on next-generation sequencing and is a big contributor to knowledge sharing in the genomics field. Other platforms exist but are often closed communities and don't contribute as much to knowledge sharing \cite{li2013ngs}.\\

Knowledge sharing is just one aspect to collaboration. Other forms of collaboration are also necessary, specifically when researchers are in the process of analysing the results of their experiments, such as sequencing data. There is often a need for synchronous collaboration where researchers discuss their findings while viewing the results. Researchers often use visualisations of the data to analyse it which quickly becomes an unfavourable task when researchers are in different geo-locations. A project called DICODE has been designed for the biomedical field in an effort to solve this synchronous collaboration problem. This project incorporates a discussion view, mind-map view, and formal view (predefined knowledge items) allowing teams to collaborate on a project in different ways \cite{karacapilidis2011facilitating}. To create a fully collaborative environment there needs to be support for communication, data sharing, and coordination all relying heavily on technical infrastructure \cite{lee2007facilitating}.\\

This literature review will be focussing on web technologies which enable collaboration such as communication and data sharing, and specifically synchronized viewing across web browsers. Synchronizing web applications across different web browsers has become a popular feature for enabling collaboration. These web applications are often referred to as groupware. A popular example is Google Docs which is an online text editor. Google Docs enables multiple users to share a single document and are consequently able to view and edit that single document simultaneously. The document synchronizes edits across each user's web browser in real-time \cite{koren2013shared}. More frameworks and libraries are being developed to simplify the integration of synchronization features into any web application \cite{ozono2012real}. The main problem areas with implementing this synchronization are conflict resolution (if two or more users edit the same area in a document the edits somehow need to be merged) and response times (how long it takes for changes to propagate to other connected users) \cite{heinrich2012enriching}. Visualisations can also be synchronized across web browsers, for example viewing a 3D model while moving it around to see it from different angles \cite{marion2012real}. The reason for this literature review is researching the feasibility of implementing synchronization of 2D visualisations, more specifically synchronizing a genome browser to help bioinformaticians analyse a genomic sequence collaboratively.  

\section{Background}
Before we look at new technologies which are most commonly used in groupware, we will look at some older technologies which were used and what their shortcomings are when it comes to implementing groupware. Synchronization requires a bidirectional data stream so that if a user (you) makes a change then that change can be pushed to the other users and if another user makes a change then that change can be pushed to you \cite{linner2012instant}.

\subsection{AJAX} 
AJAX stands for asynchronous JavaScript and XML. This technology allows requesting data from a web server using Javascript without reloading the webpage. It is widely used in websites today and it works well. The problem with AJAX is that it is not bidirectional, meaning the web server cannot push data to the client. The client always has to send a request to the web server first and then the web server subsequently responds by sending the data. This limitation can be overcome by polling the web server every interval to receive updated data. Every time a poll is made, a new connection is created with the web server. This is however very inefficient both in terms of the web server's resources and synchronization response times \cite{gutwin2011real}. 

\subsection{Comet}
Comet extends AJAX to overcome its limitation of not being bidirectional. It essentially relies on exploiting AJAX to do things it was never made for. Comet consists of three different hacks which allow bidirectional data streams. \\
\textbf{Long polling} sends a request to the web server using AJAX, but instead of receiving a response immediately, the request is only completed when the web server has new data to push to the client. This means the connection between the client and server may be kept open for much longer than a normal AJAX request would. After the client receives data that connection is closed and a new request is initiated. When data is being rapidly updated this becomes as inefficient as normal AJAX would've been.\\
\textbf{XHR Multipart Streaming} also uses AJAX but exploits a content type called multipart essentially allowing a connection to stay open between multiple messages. Multipart officially exists to allow large files to be sent in chunks. This often uses up much more memory on the client's device than necessary as it keeps the chunks in memory until the connection is closed. This can however be rectified by periodically closing the connection and creating a new request.\\
\textbf{XHR Iframe Streaming} makes use of a hidden iframe which connects to a web server and streams Javascript dynamically. The iframe is then processed to extract the data from the Javascript. This technique also results in memory problems just like multipart streaming above \cite{gutwin2011real}.\\
These exploits were used when there was no alternative besides using plugins and probably resulted in many unforeseen bugs.

\subsection{Plugins}
The final alternative to using AJAX or Comet used to be plugins. Plugins are essentially embedded applications into a web page. These applications were written as either Java applets, Flash plugins or Silverlight plugins. This allowed developers to access some native operating system APIs which weren't supported by web browsers themselves. In particular, plugins are able to create TCP socket connections with servers, which allowed access to a bidirectional data stream. This was probably the best solution as it allowed low latency persistent connections without using exploits like Comet. Plugins had other problems though, the first of which being that a client needs to install the plugin before being able to use the web application. Security was also a major problem with plugins, so much so that Apple refused to include Flash support in their iOS browser \cite{gutwin2011real}.\\\\


\section{Discussion}
Luckily for us we no longer have to deal with these old technologies in order to have access to a bidirectional data stream. Two new technologies have been created since then, namely WebSocket and WebRTC.



\section{Conclusions}
sd
\appendix
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{rtc}

\end{document}

